### 2022-05-20 paper.Rmd
some restructuring
writing

source("misc-functions.R")

@@

Anticlustering is a collection of methods to partition elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998]. Oftentimes, both goals coincide: several anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990]. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006]. Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].

@@

Table 1 displays the global simulation results aggregated across all conditions. Standard k-means is best at minimizing discrepancy with regard to means, but disregards all other distribution characteristics. K-plus anticlustering also addresses $\Delta \mathit{M}$ quite well and at the same time tends to minimize discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$). With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods---including diversity anticlustering.

As Figure 1 shows, with increasing $N$, k-means and all k-plus methods methods converge on minimizing $\Delta \mathit{M}$. This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time. In particular, standard k-plus anticlustering is not much worse than k-means anticlustering when it comes to $\Delta \mathit{M}$, but it offers the immense benefit of also equalizing the spread of the data across groups.

While k-plus anticlustering is capable of fulfilling multiple objectives at the same time, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion. For example, k-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus. Still, every k-plus method does what it is supposed to do: k-plus skew/kurtosis is best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation is best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$. At the same time, these two k-plus objectives maintain a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.

It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though this method is specifically tailored to these criteria. Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis is even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$. However, with increasing $N$, it clearly outperforms diversity anticlustering when it comes to $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$. These results show the cost of optimizing several criteria at the same time; for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$. Increasing $N$ then facilitates to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.

Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat address all distribution characteristics. For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering gets outperformed. As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups. As far as I know, this observation has not been made previously, even though the maximum diversity problem is quite well studied in the literature.

@@

In this section I demonstrate how researchers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020]. It is shown that k-plus anticlustering yields excellent results when being applied to real data sets. The results presented here used `anticlust` version 0.5.6. The package can be installed in the R environment using the `install.packages()` command:

@@

Here, the variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence. The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**). The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section. The argument `K` describes the number of groups. The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process (see [Section 1.2](#kplus)).

@@

Table 3 illustrates another advantage of k-plus anticlustering over anticluster editing: When creating groups of unequal size, anticluster editing tends to increase the spread of the data in the largest group in comparison to the other groups. K-plus anticlustering strives for equal means and variances in all groups, regardless of group size. Unequal group sizes can be requested in anticlust by passing the different group sizes to the `K` argument:

@@

In this paper, I presented the k-plus criterion for anticlustering. K-plus anticlustering focuses on constructing groups that are highly similar to each other, operationalized by similar means as well as variances. These descriptive statistics are usually "eyeballed" by applicants of anticlustering methods when they try to assess the quality of of a partitioning; hence, a criterion that directly optimizes these measures has been needed and is now provided. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data. For each input variable, a new variable is computed as the squared deviation between the original values and the overall mean of the variable. A simulation study and examples on real norming data indicate that minimizing differences with regard to means and variances between groups is well possible using the k-plus criterion.

This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time. However, for small group sizes (small $N$ and / or large $K$) k-plus anticlustering may return slightly less-than-optimal results in the eye of the user. In this case, it is not a problem to return to k-means that only focuses on the means. As opposed to some statistical workflows where "mindlessly" trying different approaches until one approach yiels satisfactory results is bad practice, such a trial-and-error approach is unproblematic for anticlustering. There is no anticluster hacking. Users may simply choose the results they like best. As a default, it is recommended to apply out k-plus anticlustering when creating groups that aim for high between-group similarity.

For optimizing similarity in means and variances it also outperforms anticluster editing, which, to be fair, does not directly strive to do so. Unequal group sizes give another advantage for k-plus anticlustering in comparison to anticluster editing.

>>>>

@@

Anticlustering is a collection of methods to partition a set of elements into disjunct groups, with the goal of obtaining high between-group similarity or within-group heterogeneity [@brusco2019; @spath1986; @valev1998]. Oftentimes, both goals coincide: some anticlustering objectives imply that similarity between groups is maximal whenever similarity within groups is minimal [@brusco2019; @papenberg2020; @feo1990]. Anticlustering thereby reverses the logic of its better known twin---cluster analysis---which seeks homogeneity within clusters and heterogeneity between clusters [@rokach2005; @steinley2006]. Whereas a variety of different approaches exist for cluster analysis such as hierarchical methods, model-based clustering, or network approaches [see @brusco_emergent_2012, for an overview], anticlustering is usually formalized as a partitioning process that divides a set of elements into disjunct groups in such a way that an objective function---representing between-group similarity and/or within-group heterogeneity---is optimized [@papenberg2020; @spath1986; @brusco2019].

@@

Table 1 displays the global simulation results aggregated across all conditions. Standard k-means performed best at minimizing discrepancy with regard to means, but disregards all other distribution characteristics. K-plus anticlustering also addresses $\Delta \mathit{M}$ quite well and at the same time tends to minimize discrepancy with regard to the standard deviations ($\Delta \mathit{SD}$). With regard to minimizing $\Delta \mathit{SD}$, all three k-plus approaches outperform all other methods---including diversity anticlustering.  As Figure 1 shows, with increasing $N$, k-means and all k-plus methods methods converge on minimizing $\Delta \mathit{M}$. This is an important observation because it shows that including the k-plus term in the optimization process does not necessarily prevent minimizing discrepancy in means; instead, multiple objectives can be addressed at the same time.

While k-plus anticlustering is capable of addressing multiple objectives simultaneously, it is apparent that increasing the number of optimization criteria aggravates fulfilling each single criterion. K-plus skew/kurtosis and k-plus correlation are worse at minimizing discrepancy with regard to standard deviations than standard k-plus, which is slightly worse at minimizing $\Delta \mathit{M}$ than k-means anticlustering. Still, every k-plus method does what it is supposed to do: k-plus skew/kurtosis is best at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$, k-plus correlation is best at minimizing discrepancy with regard to $\Delta \mathit{Cor}$. At the same time, these two k-plus objectives maintain a comparably good level of addressing $\Delta \mathit{M}$ and $\Delta \mathit{SD}$, outperforming diversity anticlustering.

It is maybe surprising that k-plus skew/kurtosis does not seem to be much better than diversity anticlustering with regard to skew and---in particular---kurtosis, even though the method has been specifically tailored to these criteria. Figure 1 gives a more fine grained assessment of this observation by splitting the results by $N$: Surprisingly, for small $N$, k-plus skew/kurtosis is even worse than diversity anticlustering at minimizing $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$. However, with increasing $N$, it clearly outperforms diversity anticlustering. These results illustrate the cost of optimizing several criteria at the same time: for low $N$, k-plus seems skew/kurtosis to be preoccupied with fulfilling the objectives $\Delta \mathit{M}$ and $\Delta \mathit{SD}$. Increasing $N$ then facilitates to also address $\Delta \mathit{Skew}$ and $\Delta \mathit{Kurtosis}$.

Generally, it should be noted that diversity anticlustering is a good all-arounder method that tends to somewhat address all distribution characteristics. For the specific objectives that are addressed by the k-plus methods, however, diversity anticlustering gets outperformed. As as side note, apart from k-plus correlation, diversity anticlustering is the only method that equalizes correlation structure among groups. As far as I know, this observation has not been made previously, even though the maximum diversity problem has been widely studied.

@@

This section demonstrates how researchers and other data analysists can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020]. It is shown that k-plus anticlustering yields excellent results when being applied to real data sets. The results presented here used `anticlust` version 0.6.1. The package can be installed in the R environment using the `install.packages()` command:

@@

Here, the variable `features` has to be a data table containing the image data: 900 rows representing images; 3 columns representing the features beauty, arousal, and valence. The accompanying OSF repository contains the fully reproducible code that includes downloading and reading the OASIS data, selecting the relevant columns, and finally storing them in the `features` variable (**TODO**). The arguments `objective = "kplus"` and `method = "local-maximum"` ensure that the k-plus criterion is optimized using the local maximum search method described in the previous section. The argument `K` describes the number of groups. The argument `standardize = TRUE` ensures that all features are standardized before the optimization starts; in particular, this option enforces that equalizing means and variances receives the same weight during the optimization process.

@@

Table 3 illustrates another advantage of k-plus anticlustering over diversity anticlustering: When creating groups of unequal size, diversity anticlustering tends to increase the spread of the data in the largest group in comparison to the other groups. K-plus anticlustering strives for equal means and variances in all groups, regardless of group size. Unequal group sizes can be requested in anticlust by passing the different group sizes to the `K` argument:

@@

In this paper, I presented the k-plus criterion for anticlustering. K-plus anticlustering focuses on dividing a data set into subgroups in such a way that these groups are similar to each other. The k-plus criterion extends the classical k-means criterion, which only represents how similar groups are with regard to their means. The k-plus family may be used to equalize the variance as well, any higher order moments such as skewness and kurtosis, as well as covariances. Interestingly, k-plus basically does not change the original k-means criterion because it can be reduced to an augmentation of the input data. A simulation study and examples on real norming data indicated that the k-plus criterion is well suited to maximize between-group similarity with regard to multiple criteria.

This work indicates that k-plus anticlustering should be the default method as a replacement of k-means anticlustering when it comes to creating similar groups, given that similarity in means and variances can usually be achieved succesfully at the same time. Depending on a user's needs, equalizing higher order moments can also be requested, but it should be noted that fulfilling several criteria at the same time may be difficult to achieve for small data sets. In any way, a user can just try out different objectives, i.e., start greedy by requesting that multiple distribution moments are equalized, and if the results for the means and standard deviations are subpar, try out fewer criteria. K-plus anticlustering was also shown to outperform diversity anticlustering with regard to the specific objectives that are included in the optimization process; diversity is still a reasonable all-arounder that strives for overall betwee-group similarity. Which method should be employed depends on a user's preferences. In doubt, several anticlustering methods can be applied sequentially and the results can be compared directly. Unlike in statistical methods, there is no problem with trying out different approaches when it comes to anticlustering.
